---
title: "Esm204_ass2"
author: "Jeremy Knox"
date: "4/27/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, warning=F, echo=F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(stargazer)
library(latex2exp)
library(finalfit)
library(janitor)
library(RColorBrewer)

gas_raw = read_csv("Gas_Data.csv")
demand = read_csv("Gas_Data.csv") %>% 
  rename(price = Price_Dollars) %>% 
  rename(qhigh = Q_High_Gallons_per_Day) %>% 
  rename(qlow = Q_Low_Gallons_per_Day)

```

# Raw Data
```{r Plot Raw Data, warning=F, echo=F, fig.height=4}

plot(demand$qhigh, demand$price, 
     xlim = c(70000, 150000), 
     xlab = "Quantity Gas (gallon)", 
     ylab = "Price ($/gallon)",
     title("Low Income Demand for Gas"))  
plot(demand$qlow, demand$price, 
     xlim = c(70000, 150000), 
     xlab = "Quantity Gas (gallon)", 
     ylab = "Price ($/gallon)",
     title("High Income Demand for Gas"))

```
#### Units: Demand measured in $/gallon. Quantity supplied measured in gallons of gasoline. Prices measured in $. Consumer/Producer surplus measured in $, as well as environmental damages.  


```{r Problem 1, include=FALSE}
# Find demand curves using linear regression:

demand_high = lm(data = demand, price ~ qhigh)
demand_low = lm(data = demand, price ~ qlow)

int_h = demand_high$coefficients[1]
q_high = abs(demand_high$coefficients[2])

int_l = demand_low$coefficients[1]
q_low = abs(demand_low$coefficients[2])

# Writing functions for demand:

d_high = function(q) {
  int_h - (q_high*q)
}

d_low = function(q) {
  int_l - (q_low*q)
}

# Demand as a function of quantity add together for aggregate demand:

d_highq = function(p){
  (int_h/q_high) - (p/q_high)
}

d_lowq = function(p) {
  (int_l/q_low) - (p/q_low)
}


# Find the slope and intercept for the aggregate demand:
agg_int = (int_h/q_high) + (int_l/q_low)
agg_slope = abs((-1/q_high) + (-1/q_low))

# Write aggregate demand function as function of price:
d_aggq = function(p) {
  agg_int - agg_slope*p
}

# Change to a function of quantity:
d_agg = function(q) {
  (agg_int/agg_slope) - ((1/agg_slope)*q)
}

# Quantity at equilibrium price of 5: 
q_sq = d_highq(5) + d_lowq(5)

# Find mpc by plugging in 5 for the AD curve:

mpc_slope = (5/q_sq)

mpc = function(q) {
  mpc_slope*q
}

mec = function(y) {
  2
}


# Consumer Surplus

# First find quantity where agg demand starts - quantity where price is equal to intercept of d_low
q_cshigh = d_highq(int_l)

# Integrate to find area under curve for the high demand through quantity where agg starts 
area_high = integrate(d_high, lower = 0, upper = q_cshigh)

# Integrate agg demand from quantity it starts through the equilibrium quantity 
area_agg = integrate(d_agg, lower = q_cshigh, upper = q_sq)

# Find total cost at equilibrium (price*quantity)
cost = 5*q_sq

# Add integrals and subtract cost  
cs_total = (area_high$value + area_agg$value) - cost


# Producer Surplus

# Integrate MPC find the area under curve through equilibrium quantity 
mpc_area = integrate(mpc, lower = 0, upper = q_sq)

# Find PS by subtracting area from cost 
ps_total = cost - mpc_area$value

# Env cost

tec = 2*q_sq
  
```

```{r Graphing Curves, warning=F, message=F, echo=F}
figure1 = ggplot(data.frame(x=c(0:400000)), aes(x=x)) + 
  stat_function(fun=d_high, geom ="line", color = "red") + 
  stat_function(fun=d_low, geom = "line", color = "blue") +
  stat_function(fun=mpc, geom = "line") +
  stat_function(fun=mec, geom = "line", color = "grey") +
  stat_function(fun=d_agg, geom = "line", color = "purple")+
  labs(x = "Quantity of Gas", y = "Price")+
  scale_x_continuous(limits =c(0,400000), expand = c(0,0))+
  scale_y_continuous(limits=c(0,25), expand=c(0,0))+
  theme_classic()
figure1
```

# 1.  
Demand (aggregate) curve under status quo is $Price = `r agg_int/agg_slope` - `r 1/agg_slope`Q$.   
Supply curve under status quo is $Price = `r mpc_slope`*Q$.   
Consumer Surplus under status quo market equalibrium is $`r cs_total`$.   
Producer surplus under status quo market equalibrium is $`r ps_total`$.   
Environmental cost under status quo is $`r tec`$.   






```{r Problem 2, include=FALSE}
# CS for high demand - area under curve from 0 to quantity at price 5 - (quantity*5)
q_high_sq = d_highq(5)
fullarea_high = integrate(d_high, lower = 0, upper = q_high_sq)
cost_high = q_high_sq * 5
cs_high = fullarea_high$value - cost_high

# CS for low demand - area under curve from 0 to quantity at price 5 - (quantity*5) 
q_low_sq = d_lowq(5)
fullarea_low = integrate(d_low, lower = 0, upper = q_low_sq)
cost_low = q_low_sq * 5
cs_low = fullarea_low$value - cost_low

cs_high_share = cs_high / (cs_high + cs_low)*100
```

# 2.  
High income consumer surplus is $`r cs_high`$.   
Low income consumer surplis is $`r cs_low`$.   
High income have $`r cs_high_share` \%$ of the consumer surplus. 




```{r Problem 3, include=FALSE}

# New equation for MPC with tax:

mpc_tax0.5 = function(q) {
  0.50 + mpc_slope*q
}

# Find quantity where the MPC and aggregate demand intersect
q_tax0.5 = ((agg_int/agg_slope)-0.50)/((1/agg_slope)+mpc_slope)

# Price from aggregate demand with the new quantity

p_tax0.5 = d_agg(q_tax0.5)

# CS for high demand = area under curve from 0 to quantity at new price - (new quantity*new price)
q_high_tax0.5 = d_highq(p_tax0.5)
area_high_tax0.5 = integrate(d_high, lower = 0, upper = q_high_tax0.5)
cost_high_tax0.5 = q_high_tax0.5 * p_tax0.5
cs_high_tax0.5 = area_high_tax0.5$value - cost_high_tax0.5


# CS for low demand = area under curve from 0 to quantity at new price - (new quantity*new price) 
q_low_tax0.5 = d_lowq(p_tax0.5)
area_low_tax0.5 = integrate(d_low, lower = 0, upper = q_low_tax0.5)
cost_low_tax0.5 = q_low_tax0.5 * p_tax0.5
cs_low_tax0.5 = area_low_tax0.5$value - cost_low_tax0.5


# PS = (new quantity*price at MPC) - area under curve through new quantity 
mpc_area_tax0.5 = integrate(mpc_tax0.5, lower = 0, upper = q_tax0.5)
cost_tax0.5 = q_tax0.5 * p_tax0.5 
ps_tax0.5 = cost_tax0.5 - mpc_area_tax0.5$value

# Total environmental damage.
tec_tax0.5 = 2*q_tax0.5


# Revenue from taxes = (quantity with tax) * (price of the tax)

# Calculate Revenue:
tax0.5_rev = q_tax0.5 * 0.50

```

# 3.  
a. The amount of gasoline produced and sold in France is $`r q_tax0.5`$.  
b. The price of gasoline is $`r p_tax0.5`$.  
c. Welfare of “High” income consumers is $`r cs_high_tax0.5`$.  
d. Welfare of “Low” income consumers is $`r cs_low_tax0.5`$.  
e. Welfare of gas producers is $`r ps_tax0.5`$.  
f. Total environmental damage is $`r tec_tax0.5`$.  
g. Total revenue generated by the tax is $`r tax0.5_rev`$.  



# 4.  
```{r Problem 4, echo=F, warning=F}

# Tax incremenets
tax_i = seq(0,5, by = 0.5)
  
# Replace MPC intercept
mpc_tax = function(q) {
  tax_i + mpc_slope*q
}

# Price at equilibrium (status quo)
p_sq = 5
#Quantity with tax:
q_tax = ((agg_int/agg_slope)-tax_i)/((1/agg_slope)+mpc_slope)
#Price with tax:
p_tax = d_agg(q_tax)
#Price of MPC with quantity of tax:
p_mpc_tax = mpc(q_tax)
#Environmental cost of tax:
tec_tax = 2*q_tax
# Total Revenue
rev_tax = tax_i * q_tax


### Overall welfare of “High” income consumers
q_high_tax = d_highq(p_tax)

# Revenue for High Income Consumers: 
rev_highcon = rev_tax * ((q_high_tax)/q_tax)

# Define the integration function:

area_h = function(q) {
  integrate(d_high, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
high_areas = sapply(q_high_tax, area_h)

# Cost:
cost_high_tax = q_high_tax * p_tax

# Calculate CS:
cs_high_tax = (high_areas - cost_high_tax) + rev_highcon

#Output information in table:
cs_high_table = as.data.frame(cbind(tax_i, cs_high_tax)) 



### Overall welfare of “Low” income consumers
q_low_tax = d_lowq(p_tax)

# Revenue for Low Incoming Consumers:
rev_lowcon = rev_tax * ((q_low_tax)/q_tax)

# Define the integration function:

area_l = function(q) {
  integrate(d_low, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
low_areas = sapply(q_low_tax, area_l)

# Cost:
cost_low_tax = q_low_tax * p_tax

# Calculate CS:
cs_low_tax = (low_areas - cost_low_tax) + rev_lowcon - tec_tax

#Output information in table:
cs_low_table = as.data.frame(cbind(tax_i, cs_low_tax)) 


### Gas producers surplus

# Define the integration function:
area_p = function(q){
  integrate(mpc, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
prod_areas = sapply(q_tax, area_p)

# Cost using the original supply curve and price at MPC
cost_tax = q_tax * p_mpc_tax 

#Calculate PS:
ps_tax = cost_tax - prod_areas

#Output information in table:
ps_table = as.data.frame(cbind(tax_i, ps_tax)) 

tax_amount = as.data.frame(cbind(tax_i, cs_high_tax, cs_low_tax, ps_tax, rev_tax))


tax_percent = tax_amount %>%
  mutate(cs_high_change = (cs_high_tax-cs_high)/cs_high) %>% 
  mutate(cs_low_change = (cs_low_tax-cs_low)/cs_low) %>% 
  mutate(ps_change = (ps_tax-ps_total)/ps_total) 
  
tax_final = kable(tax_percent, 
                  bootstrap_options = c("striped", "bordered", "condensed"),
                  digits = 2)
tax_final

```





```{r Problem 5, include=FALSE}

# Shift intercepts to 1/2 original:
int_h_2 = int_h/2 
int_l_2 = int_l/2

# New demand functions with new intercepts 
d_high_2 = function(q) {
  int_h_2 - (q_high*q)
}
d_low_2 = function(q) {
  int_l_2 - (q_low*q)
}

# Terms of P 
d_highq_2 = function(p){
  (int_h_2/q_high) - (p/q_high)
}
d_lowq_2 = function(p) {
  (int_l_2/q_low) - (p/q_low)
}

# New aggregate demand curve 
agg_int_ev = (int_h_2/q_high) + (int_l_2/q_low)
agg_slope_ev = abs((-1/q_high) + (-1/q_low))
d_agg_2q = function(p) {
  agg_int_ev + agg_slope_ev*p
}
d_agg_2p = function(q) {
  (agg_int_ev/agg_slope_ev) - ((1/agg_slope_ev)*q)
}


# Set equal to MPC curve to find new quantities 
q_ev_low = int_l_2/(mpc_slope+q_low)

q_ev_high = int_h_2/(mpc_slope+q_high)

q_ev_agg = q_ev_high + q_ev_high

# New price of gasoline
#plug agg quantity into demand agg function to find the price 

gas_price_ev = d_agg_2p(q_ev_agg)

#double check that this makes sense

d_high_2(q_ev_high)
d_low_2(q_ev_low)

# Environmental Cost
env_cost_ev = q_ev_agg*2
```

# 5.  
a. Gas consumption by “High” income consumers is $`r q_ev_high `$
b. Gas consumption by “Low” income consumers is $`r q_ev_low `$
c. Gas price is $`r gas_price_ev `$
d. Environmental damage from gasoline is $`r env_cost_ev `$





# 6.  
```{r Problem 6, echo=FALSE}
# slope of MSC with a 2.00 tax
mpc_tax2.0 = function(q) {
  2.0 + mpc_slope*q
}

# Find quantity where the new MSC and aggregate demand intersect
q_tax2.0 = ((agg_int/agg_slope)-2.0)/((1/agg_slope)+mpc_slope)

# environmental damage

env_cost_tax2 = q_tax2.0*2

## put all the environmental damages to compare 

env_damage_table = as.data.frame(rbind(env_cost_tax2, env_cost_ev)) 

tax2_reduction = ((tec-env_cost_tax2)/tec)*100

ev_reduction = ((tec-env_cost_ev)/tec)*100

env_damage_table$Reduction = c(tax2_reduction, ev_reduction)
  
  
colnames(env_damage_table) = c("Environmental Cost ($)", "Reduction in Total Environmental Cost (%)")
rownames(env_damage_table) = c("$2.00 Gas Tax", "Demand for Electric Cars")

env_damage_final = kable(env_damage_table, digits = 2) %>% 
  kable_styling(full_width = F)

env_damage_final
```

